Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-27 manuscript_marked_up.tex
just working on 3.7 and 3.8

\noindent A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical modes, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. For a repetitive ATATATATAT sequence, we resolve a rugged dynamical landscape comprising multiple metsatable states, numerous competing hybridization/dehybridization pathways, and a spectrum of dynamical relaxations. Introduction of a G:C pair at the terminus (GATATATATC) or center (ATATGCATAT) of the sequence reduces the ruggedness of the dynamics landscape by eliminating a number of metastable states and reducing the number of competing dynamical pathways. Only by introducing a G:C pair midway between the terminus and the center to maximally disrupt the repetitive nature of the sequence (ATGATATCAT) do we recover a canonical ``all-or-nothing'' two-state model of hybridization/dehybridization with no intermediate metastable states. Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation for DNA nanotechnology applications.

@@

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of 7 mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. Each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

\textbf{Dimensionality reduction.} The featurized trajectories were then projected into a low-dimensional space in preparation for microstate clustering. The standard approach to doing so is to employ time-lagged independent components analysis (tICA) to learn a linear projection into a low-dimensional embedding that maximally preserves the kinetic variance in the data.~\citep{Perez-Hernandez2013IdentificationConstruction, Schwantes2013ImprovementsNTL9, Noe2015KineticSimulation} In this work, we instead employ state-free reversible VAMPnets (SRVs) that can be conceived of as a nonlinear version of tICA. \citep{Chen} SRVs employ neural networks to learn flexible nonlinear functions of the trajectory featurization that better approximate the slow dynamical modes of the system and have been shown to produce substantially higher resolution MSMs than those developed using tICA.\citep{Chen, Sidky2019High-ResolutionVAMPnets} SRV modes were learned independently for each system to best approximate the slow collective modes for that particular DNA sequence. SRVs were trained using the SRV package we previously developed (\url{https://github.com/hsidky/srv}) employing the default network architecture of two hidden layers each comprising 100 neurons and tanh activation functions, a learning rate of 0.001, and a batch size of 50,000. We adopted a lag time of $\tau$ = 1.2 ns as appropriately short time scale to resolve the dynamical details of the hybridization/dehybridization dynamics. \citep{Phys2011MarkovValidation} As observed by Husic and Pande, the lag time cannot be treated as a hyperparameter to be optimized via the VAMP-2 score, but must be selected as a physically-motivated choice designed to expose the dynamical motions relevant at a particular time and length scale of interest. \citep{Husic2017Note:Selection} As we shall show, this choice of lag time leads to high-resolution Markovian macrostate MSMs. We guarded against overfitting using five-fold cross-validation in which we divided the trajectory data into 50:50 training:validation splits. We observed plateau of the validation loss and no evidence of overfitting after 20 epochs of training requiring approximately 22 GPU-minutes on a single NVIDIA Tesla K80 GPU card. A VAMP-2 scoring of the cumulative kinetic variance explained as a function of number of SRV collective modes also showed no evidence of overfitting -- as would be evinced by separation of the training and validation VAMP-2 scores \citep{Sidky2019High-ResolutionVAMPnets} -- and exhibited a knee for each of the four DNA sequences after the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively (\blauw{Fig.~S1} in the \blauw{Supporting Information}), and motivating the construction of 5D, 4D, 3D, and 2D embeddings, respectively.

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionSoftware} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mix systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mix systems, respectively, by projecting the molecular simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Prinz2011OptimalDynamics, Wehmeyer2019IntroductionSoftware}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

FTIR temperature series were measured for each sequence as reported previously,~\citep{Sanstead2016} and the second SVD component along temperature was fit to a two-state all-or-nothing model to determine the fraction of intact duplex $\theta$ as a function of temperature.~\citep{Marky1987CalculatingCurves} $K_d$ and $[S]_{T_f}$ were then determined from $\theta$ at the final temperature of each T-jump measurement,
	\begin{equation}\label{e3a}
	[S]_{T_f} = [1 - \theta(T_f)]c_{tot},
	\end{equation}
	\begin{equation}\label{e3b}
	K_d(T_f) = \frac{2c_{tot}[1 - \theta(T_f)]^2}{\theta(T_f)},
	\end{equation}
where $c_{tot}$ is the total oligonucleotide concentration, which was 2 mM for all measurements.

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted the ``slow'' $k_d^\mathrm{slow}$ (10-30 $\mu$s) and ``fast'' $k_d^\mathrm{fast}$ (70-100 ns) rates estimated from the T-jump response. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

>>>>

\noindent A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical modes, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. For a repetitive ATATATATAT sequence, we resolve a rugged dynamical landscape comprising multiple metsatable states, numerous competing hybridization/dehybridization pathways, and a spectrum of dynamical relaxations. Introduction of a G:C pair at the terminus (GATATATATC) or center (ATATGCATAT) of the sequence reduces the ruggedness of the dynamics landscape by eliminating a number of metastable states and reducing the number of competing dynamical pathways. Only by introducing a G:C pair midway between the terminus and the center to maximally disrupt the repetitive nature of the sequence (ATGATATCAT) do we recover a canonical ``all-or-nothing'' two-state model of hybridization/dehybridization with no intermediate metastable states. Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization \rood{by furnishing quantitative and predictive kinetic models of the dynamical transition network between metastable states}, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation for DNA nanotechnology applications.

@@

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this work, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).} Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. \rood{With the exception of the simulation data reported in Section 3.1 where we draw comparisons against experimental data collected over a range of temperatures,} each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms}.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences \rood{at their respective melting temperatures} from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

\textbf{Dimensionality reduction.} The featurized trajectories were then projected into a low-dimensional space in preparation for microstate clustering. The standard approach to doing so is to employ time-lagged independent components analysis (tICA) to learn a linear projection into a low-dimensional embedding that maximally preserves the kinetic variance in the data.~\citep{Perez-Hernandez2013IdentificationConstruction, Schwantes2013ImprovementsNTL9, Noe2015KineticSimulation} In this work, we instead employ state-free reversible VAMPnets (SRVs) that can be conceived of as a nonlinear version of tICA. \citep{Chen} SRVs employ neural networks to learn flexible nonlinear functions of the trajectory featurization that better approximate the slow dynamical modes of the system and have been shown to produce substantially higher resolution MSMs than those developed using tICA.\citep{Chen, Sidky2019High-ResolutionVAMPnets} SRV modes were learned independently for each system to best approximate the slow collective modes for that particular DNA sequence. SRVs were trained using the SRV package we previously developed (\url{https://github.com/hsidky/srv}) employing the default network architecture of two hidden layers each comprising 100 neurons and tanh activation functions, a learning rate of 0.001, and a batch size of 50,000. We adopted a lag time of $\tau$ = 1.2 ns as appropriately short time scale to resolve the dynamical details of the hybridization/dehybridization dynamics. \citep{Phys2011MarkovValidation} As observed by Husic and Pande, the lag time cannot be treated as a hyperparameter to be optimized via the VAMP-2 score, but must be selected as a physically-motivated choice designed to expose the dynamical motions relevant at a particular time and length scale of interest. \citep{Husic2017Note:Selection} As we shall show, this choice of lag time leads to high-resolution Markovian macrostate MSMs. We guarded against overfitting using five-fold cross-validation in which we \rood{constructed five random partitions of the 40 independent simulation trajectories into a training set of 20 trajectories and a validation set of 20 trajectories}.\citep{Sidky2019High-ResolutionVAMPnets} We observed plateau of the validation loss and no evidence of overfitting after 20 epochs of training requiring approximately 22 GPU-minutes on a single NVIDIA Tesla K80 GPU card. A VAMP-2 scoring of the cumulative kinetic variance explained as a function of number of SRV collective modes also showed no evidence of overfitting -- as would be evinced by separation of the training and validation VAMP-2 scores \citep{Sidky2019High-ResolutionVAMPnets} -- and exhibited a knee for each of the four DNA sequences after the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively (\blauw{Fig.~S1} in the \blauw{Supporting Information}), and motivating the construction of 5D, 4D, 3D, and 2D embeddings, respectively.

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were coarsened into our terminal macrostate MSM.  \rood{Since we are interested in the construction of models of the equilibrium kinetics, we explicitly enforce detailed balance in the construction of the MSM. \citep{Wehmeyer2019IntroductionSoftware}} The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionSoftware} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mix systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mix systems, respectively, by projecting the molecular simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Prinz2011OptimalDynamics, Wehmeyer2019IntroductionSoftware}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

FTIR temperature series were measured for each sequence as reported previously,~\citep{Sanstead2016} and the second SVD component along temperature was fit to a \rood{two-state model} to determine the fraction of intact duplex $\theta$ as a function of temperature.~\citep{Marky1987CalculatingCurves} $K_d$ and $[S]_{T_f}$ were then determined from $\theta$ at the final temperature of each T-jump measurement,
	\begin{equation}\label{e3a}
	[S]_{T_f} = [1 - \theta(T_f)]c_{tot},
	\end{equation}
	\begin{equation}\label{e3b}
	K_d(T_f) = \frac{2c_{tot}[1 - \theta(T_f)]^2}{\theta(T_f)},
	\end{equation}
where $c_{tot}$ is the total oligonucleotide concentration, which was 2 mM for all measurements.

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted \rood{the ``slow'' $k_d^\mathrm{slow}$  and ``fast'' $k_d^\mathrm{fast}$ rates, corresponding to processes on the 10-30 $\mu$s and 70-100 ns timescales, respectively}. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to correct our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations -- $f_\mathrm{hybridized}$ = $f_\mathrm{hybridized}^\mathrm{sim}$ -- to the bulk corrected result -- $f$ = $f_\mathrm{hybridized}$  -- using the following relations:}

    \begin{equation}\label{}
    \phi = \frac{\mathrm{f_\mathrm{hybridized}^\mathrm{sim}}}{\mathrm{1 - f_\mathrm{hybridized}^\mathrm{sim}}}
    \end{equation}

    \begin{equation}\label{}
	f_\mathrm{hybridized} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
	\end{equation}

Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_{\mathrm{unfrayed}}(t) = \exp(-k_d^\mathrm{fast}t)$.
